---
title: "Statistics: Chapter 3"
author: Abdullah Al Mahmud
date: \today
output: 
  revealjs::revealjs_presentation:
    theme: sky
    reveal_plugins: ["notes", "chalkboard", "menu"]
    self_contained: false
    css: css/styles.css
    includes:
      in_header: css/header_stat.html
---

# Chapter 03 (Central Tendency)

## What is Central Tendency?

Why needed?

- Summary
- Comparison
- A value to represent all

## Criteria for a Good Measure of Central Tendency

- Well-defined

> - Understandable
> - Considers all values
> - Suitable for further analysis
> - Not affected by sample fluctuation

## Measures (Averages)

- Arithmetic Mean (AM)
- Geometric Mean (GM)
- Harmonic Mean (HM)
- Median 
- Mode
- Partition Values (Quartiles, Deciles, Percentiles etc.)

## AM

$AM=\bar x=\frac{\sum x}{n}$

If there are frequencies or weights

$\bar x=\frac{\sum f_i x_i}{\sum f_i} \space or \space \frac{\sum w_i x_i}{w_i=n}$

## Find AM 

```{r, echo=FALSE}
set.seed(100)
x <- sample(60,10)
x
```

<div onclick="klikaj('rad1')">Answer</div>
<div id="rad1" style="visibility: hidden">
```{r, echo = FALSE}
mean(x)
```
</div>
<script>
function klikaj(i) {
    document.getElementById(i).style.visibility='visible';
}
</script>

Find AM: 2, 2, 3, 3, 5, 5, 5, 8, 8, 9

There are 2 ways. 

$\bar x = \frac{2+2+...+9}{10}$
<div onclick="klikaj('rad2')">Answer</div>
<div id="rad2" style="visibility: hidden">
```{r, echo=FALSE}
x <- c(2, 2, 3, 3, 5, 5, 5, 8, 8, 9)
mean(x)
```
</div>
<script>
function klikaj(i) {
    document.getElementById(i).style.visibility='visible';
}
</script>

## Mean Using Frequency

For grouped data

| Working hours<br>(x) | Employee<br>(f) | fx |
|:-:|:-:|:-:|
| 2 | 2 | 4 |
| 3 | 2 | 6 |
| 5 | 3 | 15 |
| 8 | 2 | 10 |
| 9 | 1 | 9 |
|  | $\sum f =10$ | $\sum fx = 50$ |

$\therefore \bar x = \frac{\sum fx}{\sum f} =\frac{50}{10}=5$

## Freuency vs Weight

<div class="container">
<div class="col">
Suppose, different judges give different scores, but not all evaluation has same weight. 

| Judge | Rating<br>(x) | Weight<br>(w) | wx |
|:-:|:-:|:-:|:-:|
| 1 | 8 | 2 | 16 |
| 2 | 7 | 3 | 21 |
| 3 | 4 | 5 | 20 |
| 4 | 5 | 1 | 5 |
| 5 | 7 | 3 | 21 |
|  |  | $\sum w = 14$ | $\sum wx = 83$ |
</div>
<div class="col">
$\therefore \bar x = \frac{\sum wx}{\sum w}$


<div onclick="klikaj('rad3')">Answer</div>
<div id="rad3" style="visibility: hidden">
$\frac{83}{14}$
```{r, echo=FALSE}
round(83/14, 2)
```
</div>
<script>
function klikaj(i) {
    document.getElementById(i).style.visibility='visible';
}
</script>
</div>
</div>

## Shortcut Method for AM

<div class="container">
<div class="col">
Calculate the mean in a smart way

```{r, echo=FALSE}
set.seed(100)
x <- sample(1000:1050, 6)
x
```
</div>
<div class="col">
<div onclick="klikaj('rad4')">How?</div>
<div id="rad4" style="visibility: hidden">
<small>Subtract a number from all, say 1020</small>
```{r, echo=F}
y <- x-1020
paste0("y = ", y)
```

```{r, echo=F}
paste0("Mean of y is ", mean(y)) 
paste0("Mean of x is ", mean(y)+1020)
```
</div>
<script>
function klikaj(i) {
    document.getElementById(i).style.visibility='visible';
}
</script> 

</div>
</div>

## Comparison of Measures