---
title: "Statistics: Chapter 3"
author: Abdullah Al Mahmud
date: \today
output: 
  revealjs::revealjs_presentation:
    theme: solarized
    reveal_plugins: ["notes", "search", "chalkboard", "menu"]
    self_contained: false
    css: css/styles.css
    includes:
      in_header: css/header_stat.html
---

# Chapter 03 (Central Tendency)

## What is Central Tendency?

Why needed?

- Summary
- Comparison
- A value to represent all

## Criteria for a Good Measure of Central Tendency

- Well-defined

> - Understandable
> - Considers all values
> - Suitable for further analysis
> - Not affected by sample fluctuation

## Measures (Averages)

- Arithmetic Mean (AM)
- Geometric Mean (GM)
- Harmonic Mean (HM)
- Median 
- Mode
- Partition Values (Quartiles, Deciles, Percentiles etc.)

## AM

$AM=\bar x=\frac{\sum x}{n}$

If there are frequencies or weights

$\bar x=\frac{\sum f_i x_i}{\sum f_i} \space or \space \frac{\sum w_i x_i}{w_i=n}$

## Find AM 

```{r, echo=FALSE}
set.seed(100)
x <- sample(60,10)
x
```

<div onclick="klikaj('rad1')">Answer</div>
<div id="rad1" style="visibility: hidden">
```{r, echo = FALSE}
mean(x)
```
</div>
<script>
function klikaj(i) {
    document.getElementById(i).style.visibility='visible';
}
</script>

Find AM: 2, 2, 3, 3, 5, 5, 5, 8, 8, 9

There are 2 ways. 

$\bar x = \frac{2+2+...+9}{10}$
```{r, echo=FALSE}
x <- c(2, 2, 3, 3, 5, 5, 5, 8, 8, 9)
mean(x)
```