---
title: "Measures of Dispersion"
author: Abdullah Al Mahmud
date: \today
output: 
  revealjs::revealjs_presentation:
    theme: sky
    reveal_plugins: ["menu", "chalkboard", "zoom"]
    self_contained: false
    css: css/styles.css
    includes:
      in_header: css/header_stat.html
---
# Concept of dispersion

## What is Dispersion

## Why?

To

- estimate reliability
- compare variability of several sets of data
- enable further analysis
- have a perception of probability

## Criteria of A Good Measure

- Well-defined
- Easy to understand and compute
- Based on all observations
- Suitable for for further mathematical/statistical analysis
- Less affected by sample fluctuation
- Unaffected by outliers

## Types of Measure

- Absolute
- Relative


## Absolute Measure

```{r, echo=F}
library(DiagrammeR)
mermaid("
graph LR
Absolute-->Range
Absolute-->Mean_Deviation
Mean_Deviation-->About_Mean
Absolute-->Stanard_Deviation
Absolute-->Quartile_deviation
Mean_Deviation-->About_Median
Mean_Deviation-->About_Mode
")
```

## Relative Measure

```{r, echo=F}
mermaid("
graph LR
Relative-->Coefficient
Coefficient-->of_Range
Coefficient-->of_Mean_Deviation
Coefficient-->of_Varniation
Coefficient-->of_Quartile_Deviation
")
```