---
title: ''
output:
  pdf_document: default
  html_document: default
  word_document: default
---
\begin{center}
\begin{large}
Pabna Cadet College
\end{large}

Solution to 3rd Tutorial Examination 

Subject: Mathematics

Class: VII
\end{center}

Time: 40 minutes
\hfill
Full Marks: 20

\textbf{Answer all the questions}

1. 

Given below are the daily income and expense of ten workers.

| Income(x)        | Expense (y)   | 
| ------------- |:-------------:| 
| 120     | 80|
| 130     | 120     |   


From above data, 

i. Find $$\sum_{i=1}^{2}x_iy_i \space and \space (\sum_{i=1}^{2}x_i)(\sum_{i=1}^{2}y_j)$$

ii. Are they equal? 
iii. Find $$\sum_{i=1}^{2} \sum_{j=1}^{2}x_iy_j$$

Which of (i) is equal to it? 

iv. Find $$\sum_{i=1}^{2} \sum_{j=1}^{2}(x_i-y_j)$$

*Solution:*

i. Table for calculation 


|   Income (x)  |  Expense (y)  |        xy        |
|:-------------:|:-------------:|:----------------:|
|      120      |       80      |       9600       |
|      130      |      120      |       15600      |
| $\sum x =250$ | $\sum y =200$ | $\sum xy =25200$ |

Therefore, $$\sum_{i=1}^{2}x_iy_i=25200$$

$$(\sum_{i=1}^{2}x_i)(\sum_{i=1}^{2}y_j)=250 \times 200=50,000$$
ii. So they are not equal. 

iii. 
\begin{eqnarray} 
\sum_{i=1}^{2} \sum_{j=1}^{2}x_iy_j \nonumber \\
&=&x_1(y_1+y_2)+x_2(y_1+y_2) \nonumber \\
&=& 120(80+120)+130(80+120) \nonumber \\
&=& 50,000 \nonumber \\ 
\end{eqnarray} 

Thus, this equal to the 2nd part of (i).

iv. 

\begin{eqnarray} 
\sum_{i=1}^{2} \sum_{j=1}^{2}x_iy_j \nonumber \\
&=&(x_1-y_1+x_1-y_2)+(x_2-y_1+x_2-y_2) \nonumber \\
&=& (120-80+120-120)+(130-80+130-120) \nonumber \\
&=& 100 \nonumber \\ 
\end{eqnarray}


*Solution using R code:*

**Download R from here: https://cran.r-project.org/**

```{r}
# Lines starting with hashes (#) are comments, not executable commands. 
x <- c(120,130)
y <- c(80,120)
df <- expand.grid(x,y)
df
# Answer of i
sum(x*y)
sum(x)*sum(y)
# Answer of iii
sum(df$Var1*df$Var2)
# Answer of iv
sum(df$Var1-df$Var2)
```

## 2.i. Stem and Leaf plot

 **Do not make a table; follow this format**
 
```{r}
data <- c(16, 26, 12, 10, 27, 30, 14,  1, 25, 20)
stem(data)
```

## 2.ii. Ogive

First let's draw a histogram (NOT required; only optional)
```{r}
h <- hist(data)
```

Now the Ogive-

```{r}
intervals <- seq(from = min(h$breaks), to = max(h$breaks), by = h$breaks[2] - h$breaks[1])
intervals <- c(0, intervals[-1])
# Cumulative sums
cf = c(0, cumsum(h$counts))
plot(intervals, cf, type = "b", col = "blue", pch = 20)
grid()
```

\begin{center}
---Good Luck---
\end{center}
