---
title: "Measures of Dispersion"
author: Abdullah Al Mahmud
date: \today
output: 
  revealjs::revealjs_presentation:
    theme: sky
    reveal_plugins: ["chalkboard", "zoom", "menu"]
    self_contained: false
    lib_dir: libs
    css: css/styles.css
    includes:
      in_header: css/header.html
---

# Concept of dispersion

## What is Dispersion? {data-background=white}


<div class="container">
<div class="col">
```{r, echo=F, fig.width=6, fig.height=6}
par(bg="white")
source('../r/sketch_circle.R')
sketch_circle()
``` 
</div>
<div class="col">
</br>
</br>

- Which plot has the largest dispersion?
- Which plot has the smallest dispersion?

</div>
</div>

## Why?

To

- estimate reliability
- compare variability of several sets of data
- enable further analysis
- have a perception of probability

## Criteria of A Good Measure

- Well-defined
- Easy to understand and compute
- Based on all observations
- Suitable for for further mathematical/statistical analysis
- Less affected by sample fluctuation
- Unaffected by outliers

## Types of Measure{data-background=white}

```{r, echo=FALSE}
DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = TB]

node [shape = rectangle, style = filled, fillcolor = white, fontcolor = blue]
msr [label = 'Measures of Dispersion']
abs [label = 'Absolute Measure']
rel [label = 'Relative Measure']
md [label =  'Mean \n Deviation']
rng [label = 'Range']
sd [label = 'Stanard Deviation']
qd [label= 'Quartile deviation']
abt [label= 'About']
coef [label ='Coefficient of']
var [label ='Variance']
med [label ='Median']
mean [label = 'Mean']
mod [label ='Mode']


{msr}  -> {abs, rel} 
abs -> rng
abs -> {md, sd, qd}
rel -> coef
coef -> {rng, md, qd, var}
md -> abt
abt -> {mean, med, mod}
}")
```

## Range

<div class="container">
<div class="col">
Ungrouped Data
Range, $R = X_H - X_L$
</div>
<div class="col">
Grouped Data
Range, $R = R_u - R_l$

$L_u$ = Upper boundary of the highest class

$L_l$ = Lower boundary of the lowest class

</div>
</div>

$X=$ `r set.seed(300);(x=sample(30,15))`

> - Range = `r range(x)[2]-range(x)[1]`

## Dis(advantages) of Range

- [x] Simple & quick
- [ ] <span style="color: red;">Influenced by outliers</span>
- [ ] <span style="color: red;">Not suitable for for further mathematical analysis</span>
- [ ] <span style="color: red;">Cannot be computed for open-ended distribution</span>

## Mean Deviation (MD)

<div class="container">
<div class="col">
Ungrouped Data

$MD(k)=\frac{\sum_{i=1}^n |x_i-k|} n$, not $MD(k)=\frac{|\sum_{i=1}^n(x_i-k)|} n$; Beware!

</div>
<div class="col">
Grouped Data
$MD(k)=\frac{\sum_{i=1}^n f_i|x_i-k|} n$ 
</div>
</div>
 k =
\begin{cases}
\bar x, & \text{for MD about mean} \\
\tilde x, & \text{for MD about median} \\
Mo, & \text{for MD about Mode}
\end{cases}

## Compute MD 

Compute about mean, median and mode 

> - X = `r x`

## Dis(advantages) of MD

- [x] Simple
- [x] Unaffected by outliers 
- [ ] <span style="color: red;">Useful for symmetric data only</span>
- [ ] <span style="color: red;">Not suitable for for further mathematical analysis</span>

## Variance and SD


<div class="container">
<div class="col">
### Variance 

$$\sigma ^2 = \sum_{i=1}^n \frac{(x_i-\bar x)^2}{n}$$

$\quad$ = $\sum \frac{x_i^2}{n}-(\frac{\sum x_i}{n})^2$

$\quad$ =  Mean of square - square of mean
</div>
<div class="col">
### *Standard Deviation* ($\sigma$)

Positive square root of variance

> - $$\sigma  = \sqrt{\sum_{i=1}^n \frac{(x_i-\bar x)^2}{n}}$$
> - For grouped data? 
</div>
</div>

## Variance Estimation

| $x$ | $x^2$ |
|------:|:-----:|
|   12  |  144  |  
|  11  |  121 |  
|  3   |    9 |  
|$\sum x = 26$ |$\sum x^2 = 274$|

$\sigma^2$=Mean of square - square of mean

$\quad$ =$\bar {x^2}-{\bar x}^2$

> - $\therefore \sigma^2 = \frac{274}{3}-(\frac{26}{3})^2 =$ `r (x=c(3,11,12)); round(var(x),3)`
> - $\sigma =$ `r round(sd(x),3)`

## Thoughts on Variance?

### <span style="color: blue;">Think</span>

- What is the unit of variance
- What is the unit of sd?
- Why is variance determined before sd? 
- What do variance and sd mean?

## Dis(advantages) of SD

- [x] Well-defined
- [x] Less affected by sample fluctuation
- [x] Useful for further mathematical analysis
- [x] Measures consistency
- [ ] <span style="color: red;">Difficult to compute</span>
- [ ] <span style="color: red;">Affected by outliers</span>

## Variance Example

X = `r set.seed(30); (v=sample(20:35,8))`

Find variance and stand deviation

> - Variance, $\sigma^2=$ `r round(var(v),3)`
> - SD, $\sigma$ = `r round(sd(v),3)`

## Quartile Deviation(QD)

<div class="container">
<div class="col">
$QD=\frac{Q_3-Q_1} 2$

$Q_3-Q_1$ is called Interquartile range. 
</div>
<div class="col">
### Dis(advantages)

- [x] Simple
- [x] Unaffected by outliers
- [x] Can work with open-ended distribution
- [ ] <span style="color: red;">Not based on all observations</span>
- [ ] <span style="color: red;">Not suitable for for further mathematical analysis</span>
</div>
</div>

## Coefficients 
<small>
- Coefficient of Range,$CR=\frac{X_H-X_L}{X_H+X_L}\times 100= \frac{\text{Range}}{X_H+X_L}\times 100=\frac{X_u-X_l}{X_u+X_l}\times 100$ (grouped)

> - Coefficient of Mean Deviation, $CMD=\frac{MD(\bar x)}{\bar x} \times 100$ (about mean, median and mode similarly)
> - Coefficient of Variance, $CV=\frac{\sigma}{\bar x} \times 100$
> - Coefficient of Quartile Deviation, $CQD=\frac{Q_3-Q_1}{Q_3+Q_1} \times 100$

</small>

## Compute Coefficients

X = `r v`

```{r, echo=F}
source('../r/dispersion.R')
```

> - CR = `r cr(v)`
> - $CMD(\bar x)$ = `r round(cmdm(v),3)`
> - CV = `r round(cv(v),3)`
> - CQD = `r round(cqd(v),3)`

## SD vs CV

<div class="container">
<div class="col">
- 60 students earned GPA 5 from VNC
- 45 students earned GPA 5 from Udayan college. 

 Which college is better?
</div>
<div class="col">
- Average batting score of Soumya = 35
- Average batting score of Mushfiq = 34

Who is better?
</div>
</div>

## Use of CV

- For estimating skewness, correlation etc.
- A measure of consistency/performance

## Properties of SD

> - Depends on scale, but not on origin (prove)
> - For two unequal observations, $SD=\frac R 2$; where $R = Range$
> - $MD \le SD$
> - For n positive observations, $\bar X \sqrt{n-1}\ge \sigma$
> - For the first n natural numbers, $\sigma = \sqrt{\frac{n^2-1}{12}}$
> - For two positive observations, $AM>SD$
> - For n number of unequal values, $\sigma \lt R$
> - $SD \lt \text{Root Mean Square Deviation}$

## Is $\sigma^2$ alwyas $\gt \sigma$

Exceptions

We know, $\sigma=\sqrt \sigma^2$ (N:B: $-2=-\sqrt 4; -2\ne \sqrt 4$)

> - If $\sigma^2=1, \sigma =1$
> - If $\sigma^2 \lt 1, \sigma ^2 < \sigma$
> - Example: $\sigma^2=0.05, \sigma = \sqrt{0.05} = `r sqrt(0.05)`$
 
## CV is a Pure Number

- No Unit
- Absolute number

## Minimum Value of $\sigma$

<small>

$\sigma \ge 0$

$\therefore$ Least value of $\sigma$ is 0. 

$\Rightarrow  \frac{\sum(x_i-\bar x)^2}{n} = 0$ 

$\Rightarrow  \sum(x_i-\bar x)^2 = 0$  

$\Rightarrow  (x_1-\bar x)^2 + (x_2-\bar x)^2 + \cdots + (x_n-\bar x)^2= 0$

$\therefore (x_1-\bar x)^2 =0, (x_2-\bar x)^2=0,  (x_n-\bar x)^2=0$

$\Rightarrow x_1=\bar x, x_2=\bar x, x_n=\bar x$

$\Rightarrow x_1=x_2= \cdots =x_n$

$\therefore$ SD is least (i.e., 0) when all values are equal.

</small>

## Comparison with Mean 

| Subject   | Bangla | English | Mathematics | Statistics | Average |
|-----------|--------|---------|-------------|------------|---------|
| Student X | 70     | 70      | 80          | 72         | 72.5    |
| Student Y | 98     | 96      | 48          | 50         | 72.5    |

Who is better?

# Theorems

## $\sigma^2$ Origin-Scale

<small>

<div class="container">
<div class="col">
$\sigma_x^2=\frac{\sum(x_i-\bar x)^2}{n}$

Let, $d_i=\frac{x_i-a}{c}$ (a = origin, c = scale)

\begin{eqnarray} 
&\Rightarrow& x_i=a+cd_i     \nonumber \\
&\Rightarrow& \bar x = a + c \bar d \nonumber \\
\end{eqnarray}
</div>
<div class="col">
\begin{eqnarray} 
\sigma_x^2&=&\frac{\sum(x_i-\bar x)^2}{n}     \nonumber \\
&=&  \frac{\sum(a+cd_i-a-c \bar d)^2}{n}   \nonumber \\
&=&  \frac{\sum(cd_i-c \bar d)^2}{n}   \nonumber \\
&=&  \frac{c^2\sum(d_i-\bar d)^2}{n}   \nonumber \\
&=&  c^2 \sigma_d^2  \nonumber \\
\end{eqnarray}

$\therefore \sigma_x^2=c^2 \sigma_d^2$ 
</div>
</div>

- Similar procedure and outcome for standard deviation

</small>

## MD, SD and Range 

<small>
 
<div class="container">
<div class="col">
$MD=SD=\frac R 2$ for $x_1\ne x_2$ (two unequal observations)

$\bar x = \frac {x_1+x_2} 2$ and $R=|x_1-x_2|$
</div>
<div class="col">
Mean Deviation, 

\begin{eqnarray} 
MD &=& \frac{\sum_{i=1}^2 |x_i-\bar x|}{2}     \nonumber \\
MD &=& \frac{\sum_{i=1}^2 |x_1-\bar x|+|x_2-\bar x|}{2}     \nonumber \\
&=& \frac{|x_1-\frac{x_1+x_2} 2|+|x_2-\frac{x_1+x_2} 2|}{2} \nonumber \\
&=& \frac{2|\frac{x_1-x_2}{2}|}{2}    \nonumber \\
&=& |x_1-x_2| \nonumber \\
&=& \frac R 2 \nonumber \\
\end{eqnarray} 
</div>
</div>

Similar process for SD; Start from SD formula
</small>

## SD and Range

<small>
For two unequal observations, $SD=\frac R 2$

\begin{eqnarray} 
SD&=&\sqrt{\frac{\sum_{i=1}^2 (x_i-\bar x)^2}{2}}     \nonumber \\
&=&  \sqrt{\frac{(x_1-\bar x)^2+(x_2-\bar x)^2}{2}}  \nonumber \\
&=&\sqrt{\frac{(x_1-\frac{x_1+x_2} 2)^2+(x_2-\frac{x_1+x_2} 2)^2}{2}} \nonumber \\
&=&\sqrt{\frac{(\frac{x_1-x_2}{2})^2+(\frac{x_2-x_1}{2})^2}{2}}\nonumber \\
&=&\sqrt{\frac{2 (\frac{x_1-x_2}{2})^2}{2}}\nonumber \\
&=&\sqrt{\frac{(x_1-x_2)^2} 2}=\frac{|x_1-x_2|}{2} = \frac R 2 \nonumber \\
\end{eqnarray}

</small>

## Variance of First n Natural Numbers

<small>

\begin{eqnarray} 
\sigma^2 &=&  \frac{\sum x_i^2}{n} - (\frac{\sum x_i}{n})^2    \nonumber \\
&=& \frac{1^2+2^2+3^2+ \cdots + n^2}{n}  - (\frac{1+2+3+ \cdots + n}{n})^2 \nonumber \\
&=& \frac{\frac{n(n+1)(2n+1)}{6}}{n} - (\frac{\frac{n(n+1)}{2}}{n})^2 \nonumber \\
&=&   \frac{(n+1)(2n+1)}{6} - (\frac{n+1}{2})^2 \nonumber \\
&=& \frac{n+1}{2} (\frac{2n+1}{3}-\frac {n+1}{2})   \nonumber \\
&=& \frac{n+1}{2} (\frac{4n+2-3n-3}{6}) =  \frac{n+1}{2}(\frac{n-1}{6}) = \frac{n^2-1}{12}  \nonumber \\
\end{eqnarray}

</small>

## Mean, SD, and CV

$\bar X \sqrt{n-1}\ge \sigma$ or $CV \lt 100 \sqrt{n-1}$

## Problem 01

Tow numbers are 10 and 20; Determine Range and CV

<div onclick="klikaj('p1')"><span>Answer</span></div>
<div id="p1" style="visibility: hidden">
Range, $R = 20-10=10$

> - $\sigma = \frac R 2 = \frac{10} 2 = 5$

> - $CV = \frac{\sigma}{\bar x} \times 100 = \frac{5}{15} \times 100 = `r round(5/15*100,2)` \%$
</div>
<script>
function klikaj(i) {
    document.getElementById(i).style.visibility='visible';
}
</script>

## Find SD & MD (3)

Find SD and MD of three observations: -3, 0, 3 

<div onclick="klikaj('3')"><span>Solution</span></div>
<div id="3" style="visibility: hidden">

\begin{eqnarray} 
\sigma&=&\sqrt{\text{Mean of Square - Square of Mean}} \nonumber \\
&=& \sqrt{\frac{9+9}{3}-0}=\sqrt{6}=`r round(sqrt(6),2)`      \nonumber \\
\end{eqnarray}
</div>
<script>
function klikaj(i) {
    document.getElementById(i).style.visibility='visible';
}
</script>

## More

## Missing Numbers for Mean and SD (11)

The mean and SD of 5 observations are 4.4 and $\sqrt{8.24}$, respectively. If three of the five observations are 1, 2, and 6, find the other two.

<div class="container">
<div class="col">
<div onclick="klikaj('11a')"><span>Solution</span></div>
<div id="11a" style="visibility: hidden">
$\frac{9+x_1+x_2}{5}=4.4$

$\therefore x_1+x_2=13$

Again, $\frac{1^2+2^2+6^2+x_1^2+2^2}{5}-4.4^2=8.24$

</div>
<script>
function klikaj(i) {
    document.getElementById(i).style.visibility='visible';
}
</script>
</div>
<div class="col">
<div onclick="klikaj('11b')"><span>More</span></div>
<div id="11b" style="visibility: hidden">
$\Rightarrow x_1^2+x_2^2=138-41=97$

$\Rightarrow (13-x_2)^2+x_2^2=97$

$\Rightarrow 169-26x_2+2x_2^2=97$

$\Rightarrow x_2^2-13x_2-36=0$

$\therefore x_2=9,4; x_1=4,9$
</div>
<script>
function klikaj(i) {
    document.getElementById(i).style.visibility='visible';
}
</script>
</div>
</div>

## Converting Series of Natural Numbers

$scale, c=\text{Common Difference}$

$origin, a = \text{Firts observation - c}$

<div onclick="klikaj('orsc')"><span>Example</span></div>
<div id="orsc" style="visibility: hidden">
$X = 27, 33, 39, \cdots, 111$

$c=33-27 = 6, a = 27-6=21$

$Y = \frac{X-21}{6} = \frac{27-21}{6}=1, 2, 3, \cdots, \frac{111-21}{6}=15$
</div>
<script> 
function klikaj(i) {
    document.getElementById(i).style.visibility='visible';
}
</script>

## Thanks

<center>

Visit 

<https://lecture.statmania.info>

to see all lecture slides. 

</center>