---
title: "Time Series"
author: Abdullah Al Mahmud
date: \today
output: 
  revealjs::revealjs_presentation:
    theme: sky
    reveal_plugins: ["chalkboard", "zoom", "menu"]
    self_contained: false
    lib_dir: libs
    css: css/styles.css
    includes:
      in_header: css/header.html
---

# Time Series

## What is Time Series Data?

<div class="container">
<div class="col">
Data arranged chronologically

$Y_t = f(t); t = t_1, t_2, t_3, \cdots, t_n$
</div>
<div class="col">
Example

```{r, echo=FALSE}
set.seed(10)
ts <- data.frame(Year = 2001:2005,
                 Production = sample(20,10))
knitr::kable(head(ts, 5, booktabs = TRUE)
             )
``` 
</div>
</div>

## Components of Time Series

Four Components

> -  Trend (increase/decrease)
> - Seasonal variation
> - Cyclic variation
> - Irregular/Random variation

## Uses

> - Analyze past behavior
> - Forecasting
> - Comparison by time/place
> - Segregation of components
> - Performance measure

## Symbols

$Y_t =$ Values of series at time t

$T_t =$ Trend

$S_t =$ Seasonal 

$C_t =$ Cyclic

$R_t =$ Random/irregular

## Models

<div class="container">
<div class="col">
**Additive Model**

$Y_t = T_t + S_t + C_t + R_t$

- $C_t$ and $S_t$ can be $\pm$ve
- $R_t$ can also be $\pm$ve, but in the long run, $\sum R_t = 0$
</div>
<div class="col">
**Multiplicative Model**

> - $Y_t = T_t \times S_t \times C_t \times R_t$
> - $S_t, C_t, R_t$ refer to deviation from unit
> - $S_t$ equals unity in 1 year, $C_t$ in a cycle, and GM of $R_t$ is unity (1). 
</div>
</div>

## Comparison of Models

- Components in additive models are independent. 
- In multiplicative models, components are interwined.

# Measuring Trend

## Graphical/Free-hand Method

<div class="container">
<div class="col">

```{r, echo=FALSE}
knitr::kable(ts)
```

</div>
<div class="col">
```{r, echo=FALSE}
plot(ts, pch=19)
```
</div>
</div>

## Sem-average

<div class="container">
<div class="col">
```{r, echo=FALSE}
knitr::kable(ts)
```
</div>
<div class="col">
**Steps**

1. Separate the data into two equal parts (if odd-numbered, omit middle-most)
2. Estimate averages of each group
3. Put these two values on the scatter plot and extend
</div>
</div>

## Moving Average

| Year | Production |   3-Yearly Moving  Average  |
|:----:|:----------:|:---------------------------:|
| 2001 |     412    |              NA             |
| 2002 |     438    | ${412+438+446}\over{3}=432$ |
| 2003 |     446    |  $\frac{438+446+454}3=446$  |
| 2004 |     454    |             457             |
| 2005 |     470    |             469             |
| 2006 |     483    |  $\frac{470+483+490}3=481$  |
| 2007 |     490    |              NA             |

<div class="container">
<div class="col">

</div>
<div class="col">

</div>
</div>
